\chapter{Ebuild file format}
\label{ch:ebuild-format}

\featurelabel{bash-version} The ebuild file format is in its basic form a subset of the format of
a bash script. The interpreter is assumed to be GNU bash, version as listed in
table~\ref{tab:bash-version}, or any later version. If possible, the package manager should set
the shell's compatibility level to the exact version specified. It must ensure that any such
compatibility settings (e.\,g.\ the \t{BASH_COMPAT} variable) are not exported to external programs.

The file creation mask (\t{umask}) is set to \t{022} in the shell execution environment. It is
\emph{not} saved between phase functions but always reset to this initial value.

\featurelabel{failglob} For EAPIs listed such in table~\ref{tab:bash-version}, the \t{failglob}
option of bash is set in the global scope of ebuilds. If set, failed pattern matches during
filename expansion result in an error when the ebuild is being sourced.

Name reference variables (introduced in bash version~4.3) must not be used, except in local scope.

The file encoding must be UTF-8 with Unix-style newlines. When sourced, the ebuild must define
certain variables and functions (see chapters~\ref{ch:ebuild-vars} and~\ref{ch:ebuild-functions}
for specific information), and must not call any external programs, write anything to standard
output or standard error, or modify the state of the system in any way.

\ChangeWhenAddingAnEAPI{8}
\begin{centertable}{Bash version and options}
    \label{tab:bash-version}
    \begin{tabular}{lll}
      \toprule
      \multicolumn{1}{c}{\textbf{EAPI}} &
      \multicolumn{1}{c}{\textbf{Bash version}} &
      \multicolumn{1}{c}{\textbf{\t{failglob} in global scope?}} \\
      \midrule
      0, 1, 2, 3, 4, 5  & 3.2 & No  \\
      6, 7              & 4.2 & Yes \\
      8                 & 5.0 & Yes \\
      \bottomrule
    \end{tabular}
\end{centertable}

% vim: set filetype=tex fileencoding=utf8 et tw=100 spell spelllang=en :

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "pms"
%%% LaTeX-indent-level: 4
%%% LaTeX-item-indent: 0
%%% TeX-brace-indent-level: 4
%%% fill-column: 100
%%% End:
